[Result, but poetic]: Does LLM Use Improve Data Science Education?

### 2024-07-25 paper/paper.qmd
Finished writing perceptions and usage type visualization

@@

```{r fig-programs-pie}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Student Self-Perception of Writing Ease
#| fig-align: center

@@

ggplot(program_counts, aes(x = "", y = n, fill = program_of_study)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(title = "Distribution of Unique Programs",
       x = NULL,
       y = NULL) +
  theme_minimal() +
  scale_fill_discrete(name="Program") +
  theme(axis.text.x = element_blank(),  # Remove x-axis text
        axis.title.x = element_blank(), # Remove x-axis title
        plot.title = element_text(hjust = 1.35)
        )

@@

@@

>>>>

#Maybe a table for the years instead?

@@

```{r fig-programs-pie}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Student Distribution by Program of Study
#| fig-align: center

@@

ggplot(program_counts, aes(x = "", y = n, fill = program_of_study)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(title = "Student Distribution by Program of Study",
       x = NULL,
       y = NULL) +
  theme_minimal() +
  scale_fill_discrete(name="Program") +
  theme(axis.text.x = element_blank(),  # Remove x-axis text
        axis.title.x = element_blank(), # Remove x-axis title
        plot.title = element_text(hjust = 1.35)
        ) +
  geom_text(aes(label = n), 
            position = position_stack(vjust = 0.5))

@@

# Create a multi-bar plot for coding self-perceptions grouped by response

@@

```{r tbl-ai-familiarity}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Familiarity with AI of students
#| fig-align: center

table_data <- survey_data |>
    select(ai_familiarity)|>
    mutate(
      ai_familiarity = case_when(
        ai_familiarity == "Limited to some liberal arts, such as history courses" ~ "Somewhat familiar",
        TRUE ~ ai_familiarity
      )
    ) |>
    count(ai_familiarity) |>
    pivot_wider(
      names_from = "ai_familiarity",
      values_from = "n"
    )

kable(table_data) |>
  kable_styling(bootstrap_options = "striped", full_width = F) |>
  column_spec(1, bold = T)

```{r fig-ai_use-multibar}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: AI Use
#| fig-align: center

# Create a multi-bar plot for AI use/perceptions grouped by response

#Select first three statements
writing_perceptions <- survey_data |>
  select(
    used_ai_any_reason, 
    used_ai_in_sta302
  )

writing_perceptions_long <- writing_perceptions |>
  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
  mutate(
    Statements = case_when(
      Statements == "used_ai_any_reason" ~ "Used AI for any reason",
      Statements == "used_ai_in_sta302" ~ "Used AI in Class",
      TRUE ~ Statements
    ), 
  ) |>
  filter(!is.na(Response)) |>
  filter(Response == "No" | Response == "Yes")

# Make the plot
multi_bar <- writing_perceptions_long |>
  ggplot(aes(x = Response, fill = Statements)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(title = "Student Self-Perceptions of Coding",
       x = "Response",
       y = "Frequency") +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
      )

```{r fig-ai-appropriate-pie}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Comparison of Whether AI Use is Appropriate in School
#| fig-align: center

appropriate <- survey_data |> 
  select(ai_schoolwork_appropriate) |>
  mutate(
    ai_schoolwork_appropriate = case_when(
      ai_schoolwork_appropriate == "Appropriate" ~ "Appropriate",
      ai_schoolwork_appropriate == "Inappropriate" ~ "Inappropriate",
      TRUE ~ "It depends"
    )
  ) |>
  count(ai_schoolwork_appropriate)

appropriate

ggplot(appropriate, aes(x = "", y = n, fill = ai_schoolwork_appropriate)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Do students think AI use is appropriate in class?",
       x = NULL,
       y = NULL) +
  theme_minimal() +
  scale_fill_discrete(name="Response") +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),  # Remove x-axis text
        axis.title.x = element_blank(),
      ) +
  geom_text(aes(label = n), 
            position = position_stack(vjust = 0.5))

```{r fig-AI-usage-type-comp}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: (CHANGE) Boilerplate figure caption 
#| fig-align: center

survey_data |>
    select(
      not_used,
      used_technical_q,
      used_conversation,
      used_general_knowledge,
      used_solving_hmk,
      used_check_soln,
      used_quick_q,
      used_expl_concepts,
      used_writing_content,
      used_writing_code
    )|>
    pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
    filter(Response == 1) |>
    count(Statements) |>
    mutate(Statements = case_when(
      Statements == "not_used" ~ "Not Used",
      Statements == "used_technical_q" ~ "Technical Questions",
      Statements == "used_conversation" ~ "Fun Conversation",
      Statements == "used_general_knowledge" ~ "General Knowledge",
      Statements == "used_solving_hmk" ~ "Entirely Solving Homework",
      Statements == "used_check_soln" ~ "Checking my Solutions",
      Statements == "used_quick_q" ~ "Quick Questions",
      Statements == "used_expl_concepts" ~ "Explaining Concepts",
      Statements == "used_writing_content" ~ "Writing Paper Content",
      Statements == "used_writing_code" ~ "Writing Paper Code",
      TRUE ~ Statements
    ))|>
    ggplot(aes(x = Statements, y = reorder(Statements, n))) +
    geom_bar(fill="grey50") +
    labs(title="How Students Used ChatGPT", x="Count", y="Use Type") +
    theme(axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
          panel.background = element_blank(),
          axis.line.y=element_line(color="black"),
          axis.ticks.y=element_blank(),
          panel.grid.major.x=element_line(color="grey80"),
          panel.grid.minor=element_blank(),
          plot.title.position = "plot",
          plot.title = element_text(hjust = 0.5))

bbb
